Reproducibility of COVID-19 research papers on medRxiv

### 2021-02-15 outputs/paper/paper.Rmd
Add saved version of the checked dataset so dont have to run each time. Minor updates to paper.
Updates to .txt files, paper, scripts

---
title: "Reproducibility of COVID-19 research papers on bioRxiv and medRxiv"
subtitle: "TBD"
author: "Annie Collins^[University of Toronto.]"
thanks: "We thank CANSSI... Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "We create a dataset of all the papers published on bioRxiv and medRxiv between X and Y. We extract the text from these papers and parse them for keywords to do with the availability of data and scripts underpinning the paper. We find that X per cent of papers have X. Our paper demonstrates the need for Y."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

@@

# Data

We construct our dataset by first using the APIs of bioRxiv and medRxiv to download any paper that:

@@

@@

## First discussion point

## Second discussion point

## Third discussion point

@@

Weaknesses and next steps should also be included.

@@

>>>>

---
title: "Reproducibility of COVID-19 research papers on medRxiv"
subtitle: "TBD"
author: "Annie Collins^[University of Toronto.]"
thanks: "We thank CANSSI... Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "We create a dataset of all the papers published on bioRxiv and medRxiv between X and Y. We extract the text from these papers and parse them for keywords to do with the availability of data and scripts underpinning the paper. We find that X per cent of papers have X. Our paper demonstrates the need for Y."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(here)

knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)

```

@@

# Data & Methodology

Our primary data set consists of information extracted from the medRxiv repository combined with output from running a sample of COVID-19-related pre-prints through the Open Data Detection in Publications (ODDPub) text mining algorithm.

We constructed this data set by first creating a local copy of the medRxiv repository via the medRxiv API and then filtering for papers related to the COVID-19 pandemic through several keyword searches to create our sampling frame (n = 9,929, including only the most recent version of any given pre-print). This data includes the following variables for each pre-print in the repository: title, abstract, author(s), date posted, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, and published DOI (if the paper has since been published in a peer reviewed journal).

We then selected a random sample of these papers (n = 1,200) to check for open data and code markers using the ODDPub algorithm. This required downloading each paper as a PDF, converting the PDFs to text files, and conducting the open data and code detecting procedure to produce a results table indicating the presence of open data or open code markers in each paper (with a value of TRUE or FALSE for each marker and the relevant open data or open code statements when applicable).

Our final data set was formed by joining these two tables together via DOI to form a data set including all original, qualitative information for each pre-print alongside its open data and open code status and markers.

# Sample Summaries
```{r summary-table}
# open data/open code count summary table
med_open_data_results <- read_csv(here("outputs/data/med_open_data_results.csv"))
med_open_data_results %>% 
                    count(is_open_data, is_open_code) %>%
                    kable(col.names = c("Contains Open Data Markers", "Contains Open Code Markers", "Count"))
```

```{r open-data-published-summary}
med_open_data_results %>% 
  group_by(published) %>% count(is_open_data) %>% 
  pivot_wider(names_from = is_open_data, values_from = n) %>%
  rename("Open Data Markers" = `TRUE`, "No Open Data Markers" = `FALSE`) %>%
  kable()
```

```{r open-code-published-summary}
med_open_data_results %>% 
  group_by(published) %>% count(is_open_code) %>% 
  pivot_wider(names_from = is_open_code, values_from = n) %>%
  rename("No Open Code Markers" = `0`, "Open Code Markers" = `1`) %>%
  kable()
```

```{r monthly-papers-total}
include_graphics(here("outputs/figures/papers_posted_total.pdf"))
```

```{r monthly-papers-sample}
include_graphics(here("outputs/figures/papers_posted_sample.pdf"))
```

```{r}
include_graphics(here("outputs/figures/prop_open_per_month.pdf"))
```

@@

In this analysis we distinguish between *posting* (the event that a pre-print is uploaded to medRxiv) and *publishing* (the event that the pre-print has subsequently been published in a peer reviewed journal).

```{r}
# logit model: posting rate vs. open data, publication rate vs. open code
```

```{r}
# logit model: open data/code vs. publication status
# pub_model <- glm(published ~ is_open_data + is_open_code, family = binomial, data = med_open_data_results)
```

@@

There are many factors that impact the ability of an author to post make their data available

## First discussion point - Time
- How has open data/code shifted over the course of the pandemic?
- How has open data/code changed with publication rate?

## Second discussion point - Eventual Publication
- See weaknesses

## Third discussion point - Type of Paper
- A significantly higher proportion if

@@

- No open data/code was verified manually, all dependent on algorithm
- Publication info has high rate of false negatives (i.e. medRxiv data seems to miss a lot of papers that go on to be published)
- Want to look at geographic distribution and prevalence of open data/code - influence on open data policies and timing throughout pandemic

@@

Include info here directly from package documentation (i.e. keywords/phrases used for text parsing)?
